Do motif roles provide unique information about species risk of extinction?

### 2021-07-20 manuscript/temp_motif_persistence.tex

            A further complication when trying to understand whether relationships between motifs and probability of extinction depend upon underlying relationships between degree or trophic level and extinction 
            We also fit linear mixed effect models relating the count or proportion of each motif to degree or trophic level and a random effect of the interaction between network size and connectance using the R~\citep{R} function `lmer' from the package \emph{lmerTest}~\citep{lmerTest}.

@@

    \subsection*{Persistence did not vary strongly with global network structure}

         Network size and connectance affected persistence, but to a much smaller extent than increasing disturbance to basal resources (Figure S1, Table S1, \emph{Appendix SA}). Persistences generally decrease when networks are more connected, but the effect of network size varies depending on level of disturbance. At low disturbance, small and highly-connected webs show the lowest persistence, whereas at high disturbance, large and highly connected webs show the lowest persistence (Fig. S2, \emph{Appendix SA}).

    \subsection*{Network motif profiles varied with global network properties}

        A PERMANOVA test suggests that a network's normalized motif profile as a whole (i.e., proportions of the total made up by each motif) is related to the network's size, connectance, and the interaction between them, but unequal variation in motif profiles across global network properties may, however, mean that these PERMANOVA results are false positives (Fig. S2, \emph{Appendix SB}).
        Considering each motif separately, the proportions of apparent competition and three-species chains in a network's motif profile tended to decrease with increasing connectance while the proportion of omnivory increased with increasing connectance (Table S2, \emph{Appendix SB}). 
        Network size and the interaction between network size and connectance did not strongly affect the frequencies of most motifs (Fig.~\ref{motif_proportion_lms}). The proportion of direct competition, however, decreased with increasing connectance in small networks but showed very little change in large webs.

@@

    \subsection*{Persistence decreases with greater omnivory in a network}

        The motif profile of a network was related to the mean persistence of consumers ($F_{1,2998}$=571, $p$\textless0.001). Mean persistence was higher in networks with higher proportions of the chain, direct competition, and apparent competition motifs (Table S3, \emph{Appendix SC}). Conversely, mean persistence is lower in networks with higher proportions of the omnivory motif. These trends hold for all levels of disturbance (Table S3, \emph{Appendix SC}). Dispersion of motif profiles varies between networks with different levels of persistence ($F_{134,2865}$=2.19$\times10^{24}$, $p$=\textless0.001). For example, networks with more variable motif profiles have higher persistence ($\beta$=0.183, $p$\textless0.001).

@@

    \subsection*{Species persistence varies with motif participation}

        % See \emph{Appendix SY} for relationships between persistence and non-normalized motif participation.

        \subsubsection*{Effects of motif participation depend on disturbance}

            At low levels of disturbance, species' persistences increased with increasing proportions of the omnivory and direct competition motifs. Conversely, persistence decreased with as the  proportion of the apparent competition motif increased. Persistence was not significantly related to the proportion of the three-species chain (Figure~\ref{fig:prop_lmer_all}).

            At high levels of disturbance, persistence decreased as the proportion of the three-species chain increased, but increased as the proportion of all other motifs increased (Figure~\ref{fig:prop_lmer_all}).
            % This means that increased disturbance strengthened the effects of the three-species chain motif and changes direction of the effect from the apparent competition motif  (Table S4, \emph{Appendix SD}).
            The effect of omnivory on persistence was  positive but weaker at higher levels of disturbance, and the effect of direct competition was largely unaffected by disturbance level.

@@

        \subsubsection*{Effects of motif participation are more consistent in highly-connected networks}

            By fitting linear regressions of persistence against motif proportions for the species for each network separately, we can observe how the relationships between motifs and persistence change with the level of disturbance and global network structure. Network size had a much smaller effect than connectance (Fig. S3; \emph{Appendix SE}); we therefore focus on connectance here. 
            Networks with high connectance generally have more consistent relationships between persistence and motif participation, i.e., sharper peaks of the density distributions (Fig.~\ref{fig:density_prop_C}). 
            However, there was an interaction between disturbance and connectance for the direct competition and omnivory motifs.

            At low and medium connectances with a high proportion of the direct competition motif, there is little change in the proportion of positive slopes (\textgreater80\%) for the different levels of disturbance. 
            At high connectances, however, the proportion of positive slopes is lower overall and decreases with increasing disturbance. 
            This means that at low levels of connectance, an increased proportion of the direct competition motif increases persistence in most networks, while the opposite is true with high connectance. 
            The trend is reversed for the omnivory motif.

@@

    \subsection*{Persistence increases with degree but decreases with STL}

        Persistence varies with both species in-degree and trophic level (Fig.~\ref{fig:motifs_vs_TL_and_deg}; \emph{Appendix SF}). Increased trophic level is associated with decreasing persistence, although the rate of this decline is slower at higher levels of disturbance (Table S5, \emph{Appendix SF}).
        The interaction between in-degree and disturbance is even stronger, such as having a large number of prey is associated with higher persistence in the no disturbance scenario, but lower persistence at high disturbance.

@@

    \subsection*{Simpler measures do not fully explain effects of motifs}

        Motif participation varied with web size, connectance, and their interaction (Fig. S4, Table S6; \emph{Appendix SG}).
        Species in all networks appeared most frequently in the apparent competition motif, but the frequency of the apparent competition motif decreased with increasing network size and connectance.
        The frequency of the omnivory motif increased sharply with increasing connectance, while the frequency of the direct competition motif increased slightly with increasing network size.

        Species' motif roles also vary with their in-degree and trophic level (Fig.~\ref{fig:motifs_vs_TL_and_deg}).
        Species with higher in-degree tend to have higher proportions of their motif role made up by omnivory. Proportions of the other three motifs tend to decline with increasing degree.
        Species with higher trophic levels tend to have higher proportions of apparent competition and three-species chains.

@@

    \subsection*{Persistence varies with a species' motif role}

@@

% Results summary - network level (following Anna)
Changes in the primary producer level in food webs have profound effects on all other species in the network \citep{}. In particular, the disturbances can percolate across trophic levels as all consumers depend directly or indirectly on primary producers for their survival \citep{}. Understanding which characteristics of a network and its species influence the severity of outcomes after a disturbance is crucial, e.g. for conservation actions \citep{}. 
Here we were particularly interested in how the distribution of meso-scale structures in ecological networks are related to average network persistence when the basal level is disturbed by species loss. In order to disentangle the effects caused by the motif distribution we also analyzed global (network-wide) and local (species) properties.
%Consistent with earlier work, networks with higher connectance or species richness tend to have lower persistences~\citep{}.
A network's motif profile had relatively strong relationships to the network mean persistence, while effects from global scale structures, such as network size and connectance, were relatively small. 
However, connectance (and, to a lesser extent network size) do affect network motif profiles. Thus, global-scale properties seem to have stronger effects on network persistence by changing meso-scale properties than directly. ~\citep{prill2005dynamic, bascompte2005simple}.

% Network-level motif profiles, persistence, and relationships to S and C.
A network's mean persistence depended on which motifs were most common in the network profile. In general, persistence was lowest in networks with high frequencies of the omnivory motif and highest in networks with high frequencies of the apparent competition motif. Direct competition and three-species chain motifs had smaller effects. The effects were clearly pronounced when disturbance was low but attenuated with increased disturbance.
The negative relationship between omnivory and mean persistence is consistent with other studies showing that an omnivory motif is less likely to be stable than the other motifs considered~\citep{Borrelli2015a}.

High frequency of the apparent competition motif was correlated with an on average higher in-degree for consumers. In-degree in turn is often related to higher persistence \citep{}.

@@

% HERE ENDS MEAN PERSISTENCE AND SPECIES WISE START
As well as the mean persistence of all consumers, we also considered the species-specific persistence of each consumer in a network.
While the persistence of a particular species is likely to be affected by global-scale (size and connectance) and meso-scale (motif profiles) network structure, it also depends on a species' position in the network. 
There are many ways of measuring species' positions, and in some cases multiple measures may capture the same trend.
In our results, motif participation is significantly related to persistence, but the effects of two motifs (\emph{apparent competition} and \emph{direct competition}) might be explained more simply by in-degree and trophic level.
This is especially likely for these two motifs, since Bayesian Network modeling lacks top-down effects and therefore removes the indirect effects from these motifs.

% Getting rid of AC/DC "quickly" here so we can focus on omnivory
Participation in the apparent competition motif is strongly negatively correlated with in-degree---a high proportion of the apparent competition motif implies a low in-degree. 
Both higher in-degree and lower participation in the apparent competition motif are associated with greater persistence at low disturbance but decreased persistence at high disturbance; because the trends are so similar, it is likely that a common mechanism underlies both.
% The effect on persistence is consistent between species with a low in-degree and a high proportion of the apparent competition motif. 
%Low in-degree decreases persistence at low disturbance, and increases persistence at high disturbance, similar to the trend visible for a high participation in the apparent competition motif. 
% Similarly, the effect on species persistence is reversed for a low proportion of the apparent competition motif in which species have a high in-degree.

While the direct competition motif is not correlated with in-degree, it is strongly negatively correlated with trophic level---species with a high proportion of the direct competition motif have a low trophic levels. 
Low trophic levels are always associated with higher persistence because these species are less affected by bottom-up disturbances~\citep{}.
It then follows that species which participate in many direct competition motifs may have higher probabilities of persistence because they are also likely to have low trophic levels.

In contrast to the competition motifs, the three-species chain and omnivory motifs do have indirect effects in a Bayesian Network setting and the effects of these motifs cannot be fully explained by degree or trophic level.
For example, the three-species chain motif is negatively correlated with in-degree and positively correlated with trophic level. 
Despite this correlation, participation in the three-species chain does not interact the same way with disturbance as either simpler measure and so has a unique relationship to species persistence.
This is likely due to indirect effects of the bottom species in the chain on the top species.

The most interesting motif in which to consider indirect effects is omnivory.
While participation in the omnivory motif is positively correlated with in-degree, a high proportion of the omnivory motif is always associated with increased persistence, while the effect of in-degree varies with disturbance. 
Thus, the effect of a high proportion of the three-species chain motif and the omnivory motif does not follow the corresponding effect of in-degree and trophic level.

@@

% % Might be able to add some of this paragraph to the one above or below (especially the refs).
% The divergent effect on persistence of either having a high proportion of the three-species chain or omnivory motif can, within our Bayesian Network framework, be explained by the structure of the food web.
% When participating in a food chain motif---three-species chain or omnivory---the top species of this structure will have a higher susceptibility since disturbances amplify through the network. In our results, this is in particular evident for the three-species chain motif at higher levels of disturbance, but this also applies to species with a high proportion of the omnivory motif. While the relationship between omnivory and persistence is strongly positive at low bottom-level disturbance, the trend is reduced at high levels of disturbance. 
% This is because species participating to a high extent in omnivory motifs share the sensitivity of species participating to a high extent in three-species chain motif, and are similarly negatively affected by bottom-level disturbances. However, the top species in an omnivory motif also have the benefit of an additional prey species, located further down in the food web. This prey item can assist as a buffer, decreasing the negative effect of disturbance on the consumer species. Clearly, this additional prey is of importance, since it is more beneficial with a higher proportion of the omnivory motif for an individual species than a high proportion of the three-species chain motif. 
% Thus, the observed benefit of a high proportion of the omnivory motif can be compared to omnivores in a network feeding adaptively. Studies have demonstrated that adaptive omnivory can increase the permanence of three-species chains \citep{Fagan1997, KrivanDiehl2005, AbramsFung2010}.
% Although species in our model cannot swap prey dependent on bottom-level disturbances or lower resource densities, a species with a high proportion of omnivory will in a sense spread its risk between prey items at different trophic levels.

@@

 \cite{Cirtwill2021_inprep} analysed how species time to extinction varied depending on their network profile. They found that a high proportion of the omnivory motif resulted in a shorter time to extinction, while the other three stable motifs were associated with longer times to extinction. However, the study by Cirtwill and Wootton used a different approach using a dynamic simulation model also including top-down effects. 
    %In previous studies of the effect of omnivory motifs, the strength of the omnivorous link can potentially be important. If this link is strong, it is shown to be destabilizing (REF), while stabilizing if weaker (REF). In our Bayesian Network setting, this link is indistinguishable from any other link in the food web which partly may explain why omnivory results in increased persistence of consumer species.

    % % We can move this to the SI or delete entirely.
    % Although we performed analysis of both the effect of the absolute counts of motifs and the normalized motif role on persistence, we here focus on the normalized motif role.
    % This is because, the consistency of the relationships between persistence and the number of times a species appears in any of the possible motifs suggests that the observed effects of non-normalized motif roles are actually due to degree.
    % Species with higher degrees will with high probability appear in more motifs than species with low degrees.
    % It is possible for a species with low degree (few direct interaction partners) to participate in many motifs if it interacts with high-degree species (thereby obtaining many indirect interaction partners), but such a situation is fairly unlikely.
    % Because of this overlap between degree and absolute numbers of motifs, and because degree shows a similar trend to the motif counts [[to be confirmed]], we recommend that researchers consider degree and/or normalized motifs rather than absolute motifs.
    % When we consider the proportion of a species' role made up by each motif, the influence of degree is largely removed.
    % Given the similarly high $R^2$ values between the models using proportions of motifs and absolute counts, it does not appear that we lose much explanatory power by removing degree. 
    % Therefore, our main analysis focus on the effect of the normalized motif role on persistence.

>>>>

            A further complication when trying to understand whether relationships between motifs and probability of extinction depend upon underlying relationships between degree or trophic level and extinction 
            We also fit linear mixed effect models relating the proportion of each motif to degree or trophic level and a random effect of the interaction between network size and connectance using the R~\citep{R} function `lmer' from the package \emph{lmerTest}~\citep{lmerTest}.

@@

We are interested in i) how a the distribution of motifs in a network affect network persistence, and ii) how a species motif profile affects its persistence. We analyse this for different levels of disturbance to the basal level.

The results are organized as follows: 1) How network persistence varies with network motif distribution. 2) How network persistence varies with global structural properties. 3) How network motif distribution varies with global structural properties. 4) How species probability of persistence varies with its motif profile. 5) How species persistence varies with its characteristics such as TL and in-degree. 6) How a species motif profile depend on its characteristics. 7) How a species persistence  varies with its motif role (position in respective motif).

    \subsection*{Network persistence and level of disturbance and global network properties}
    The level of disturbance naturally had clear and profound effects on the average network persistence, such as a increased probability of extinction of basal species always decreased network persistence.

    Network size and connectance affect persistence, but to a much smaller extent than increasing disturbance (Figure S1, Table S1, \emph{Appendix SA}). Persistences generally decrease when networks are more connected, but the effect of network size varies depending on level of disturbance. At low disturbance, small and highly-connected webs show the lowest persistence, whereas at high disturbance, large and highly connected webs show the lowest persistence (Fig. S2, \emph{Appendix SA}).

    \subsection*{Network motif profile and global network properties}

        Considering each motif separately, the proportions of apparent competition and three-species chains in a network's motif profile tended to decrease with increasing connectance while the proportion of omnivory increased with increasing connectance (Table S2, \emph{Appendix SB}). Network size and the interaction between network size and connectance did not strongly affect the frequencies of most motifs (Fig.~\ref{motif_proportion_lms}).

@@

    \subsection*{Network motif profile and network persistence}

        The motif profile of a network is related to the average persistence of consumers ($F_{1,2998}$=571, $p$\textless0.001). Average persistence is higher in networks with higher proportions of the chain, direct competition, and apparent competition motifs (Table S3, \emph{Appendix SC}). Conversely, average persistence is lower in networks with higher proportions of the omnivory motif. These trends hold for all levels of disturbance (Fig.~\ref{motif_profile_persistence}, Table S3), \emph{Appendix SC}).% Network persistence varies between networks with different dispersion of motif profiles ($F_{134,2865}$=2.19$\times10^{24}$, $p$=\textless0.001). For example, networks with more variable motif profiles have higher persistence ($\beta$=0.183, $p$\textless0.001).

@@

    \subsection*{Species motif participation and species persistence}

       The effects of motif participation on a species persistence depend on level of disturbance. When disturbance is low species' persistence increase with increasing proportions of the omnivory and direct competition motifs. Conversely, a species' persistence decrease with increasing proportion of the apparent competition motif. Persistence is not significantly related to the proportion of the three-species chain motif (Figure~\ref{fig:prop_lmer_all}).

        At high levels of disturbance, species persistence decrease as the proportion of the three-species chain increase, but increase as the proportion of all other motifs increase (Figure~\ref{fig:prop_lmer_all}). This means that increased disturbance strengthened the effects of the three-species chain motif and changes direction of the effect from the apparent competition motif  (Table S4, \emph{Appendix SD}). The effect of omnivory on persistence is still positive but weaker at higher levels of disturbance, and the effect of direct competition was largely unaffected by disturbance level.

@@

        \subsubsection*{Consistency in species persistence versus motif participation}

            In order to evaluate how consistent the relationships between a species persistence and motif participation is, and how it varies with  level of disturbance and global network structure, we fit linear regressions of species persistence against motif proportions for the species for each network separately. We then analyze the density distribution of the values of the slopes. Network size had a much smaller effect than connectance (Fig. S3; \emph{Appendix SE}); we therefore focus on connectance here.

            Networks with high connectance generally have more consistent relationships between persistence and motif participation, i.e., sharper peaks of the density distributions (Fig.~\ref{fig:density_prop_C}). However, there is an interaction between disturbance and connectance for the direct competition and omnivory motifs; At low and medium connectances there is little change in the proportion of positive slopes when the proportion of  direct competition motif is high, independent of level of disturbance. At high connectances, however, the proportion of positive slopes is lower and also decreases with increasing disturbance. For the omnivory motif the trend is reversed. Additionally, the omnivory motif stands out having a wider curve when network connectance is low.

@@

    \subsection*{Species characteristics and persistence}

        A species persistence varies with both species in-degree and trophic level (Fig.~\ref{fig:motifs_vs_TL_and_deg}; \emph{Appendix SF}). Increased trophic level is generally associated with decreasing species persistence, although the rate of decline is slower at high disturbance (Table S5, \emph{Appendix SF}).
        In-degree shows an even stronger interaction with disturbance, as a having more  prey species is associated with higher persistence in the no disturbance scenario, but lower persistence at high disturbance scenario.

@@

    \subsection*{Species motif participation and network properties and species characteristics}

       Species' motif roles varied depending on the properties of the network; size, connectance, and their interaction (Fig. S4, Table S6; \emph{Appendix SG}).
        Species appeared most often in the apparent competition motif in all types of networks, but the frequency decreased with increasing network size and connectance.
        The frequency with which species participated in the omnivory motif increased sharply with increasing connectance, while the frequency of the direct competition motif increased slightly with increasing network size.

        Species' motif roles also varied with their in-degree and trophic level (Fig.~\ref{fig:motifs_vs_TL_and_deg}).
        Species with higher in-degree tend to have higher proportions of their motif role made up by omnivory. Proportions of the other three motifs tend to decline with increasing degree.
        Species with higher trophic levels tend to have higher proportions of apparent competition and three-species chains.

@@

    \subsection*{Persistence and species' positioning in motifs}

@@

% Results summary - network level (following Anna)
Changes in the basal level in food webs have profound effects on all other species in the network \citep{}, as all consumers depend directly or indirectly on primary producers for their survival \citep{}.It has repeatedly been shown that global network properties are important for how well a network can handle disturbances, such as loss of species \citep{Eklof2006, Dunne2002}. At the same time, from a conservation perspective it is often more relevant to look at specific species to which actions can be targeted \cite{}. Nevertheless, since species do exist in a network their interactions, direct as well as indirect, will affect how they respond to a disturbance. Motifs are the basic building blocks in networks and as such represents a bridge between network-wide (global) and species-specific (local) properties. Here we were particularly interested in how the distribution of these meso-scale structures in ecological networks are related to network and species persistence when the plant community is affected by species loss.

A network's motif profile had relatively strong relationships to the network average persistence, compared to global structures, such as network size and connectance, whose effects were smaller. However, connectance (and, to a lesser extent network size) do affect network which motifs that are most prominent. Thus, global-scale properties have stronger effects on network persistence by changing meso-scale properties than directly which indicates that network motifs do provide additional information.

% Network-level persistence
 In general, average network persistence was lowest in networks with high frequencies of the omnivory motif and highest in networks with high frequencies of the apparent competition motif. Direct competition and three-species chain motifs both showed positive but weaker relationships. The level of disturbance naturally affected the level of persistence, such as increased disturbance decreased average persistence, but trends were largely unaffected.

% About fraction of omnivory and apparent competition in the full network 
The negative relationship between omnivory and network persistence is consistent with other studies showing that an omnivory motif is less likely to be stable than the other motifs considered~\citep{Borrelli2015a}.

High frequency of the apparent competition in a network motif was correlated with an on average higher in-degree for consumers. In-degree has repeatedly been suggested to be beneficial for a species persistence  \citep{}.

@@

% Species level persistence
% What we want to discuss here: 1) Apparent competition change direction depending on disturbance. 2) High disturbance strengthen the effect of chain. 3) Omnivory most positive in low disturbance scenarios.

While the persistence of a particular species is likely to be affected by global-scale (size and connectance) and meso-scale (motif profiles) network structure, it also depends on what characterize a species position in the network., such as trophic level, number of interactions and which motifs the species participate in.

In our results, motif participation is significantly related to persistence, but some of the effects can to some extent be explained by local properties such as in-degree and trophic level. This is especially likely for the apparent competition and direct competition motifs, since in Bayesian Network framework lacks top-down effects and therefore the indirect effects in these motifs are not relevant.

Participation in the apparent competition motif is strongly negatively correlated with in-degree --a high proportion of the apparent competition motif implies a low in-degree. Both higher in-degree and lower participation in the apparent competition motif are associated with greater persistence at low disturbance but decreased persistence at high disturbance. This suggests that in-degree is in fact the factor driving the positive relationship between proportion of apparent competition motif and species persistence. Interestingly though, the relationship changes from positive to negative with increasing level of disturbance. When disturbance is high having many prey species acts as an insurance, while at high disturbance connections to many prey species is in fact negative as the consumer receives the negative impact from several sources.

However, the coupling between a species proportion of motif participation and degree is neither evident or always true. For example, it is possible for a species with low degree to participate in many motifs if it interacts with high-degree species and thereby obtain many indirect interaction partners. This is one of the strengths with analyses of motifs that it considers local structures additionally include indirect connections one link away.

A species participation in the direct competition motif is only weakly correlated with in-degree, but is instead strongly negatively correlated with trophic level---species with a high proportion of the direct competition motif are more likely to have low trophic levels. Low trophic levels are naturally associated with higher persistence in our analyzes  as they receive less negative impact from the bottom-up disturbances ~\citep{Eklof2013}. The clear positive relationship between proportion of direct competition motifs and species persistence is therefore likely due to the tendency for species at low trophic levels to have a higher proportion of that motif. Therefore, contrary to the apparent comptition motif, we do not observe a change in the relationship with changed disturbance.

The three species chain motif have a nearly neutral relationship with persistence when disturbance is low, but a clearly negative relationship when disturbance is high. This can partly be explained with the positive relationship between trophic level and a species proportion on the chain motif. In the three species chain motif the species are on average more often dependent on two trophic levels (if the focal species possesses the top-position in the motif), and as such affected negatively from two trophic levels. A high level of disturbance (high probability for a basal species to go extinct) will then sharply decrease persistence for the top species.

%In contrast to the two competition motifs, the three-species chain and omnivory motifs have a less evident relationship to species in-degree or trophic level. For example, the three-species chain motif is negatively correlated with in-degree and positively correlated with trophic level. Despite this correlation, participation in the three-species chain does not interact the same way with disturbance as either of the simpler measure and so has a unique relationship to species persistence. This is likely due to indirect effects of the bottom species in the chain on the top species.

The perhaps most interesting motif is omnivory. Omnivory shows a reversed relationship with disturbance level compared to the three species chain; at low disturbance a high proportion of omnivory is positive for persistence whereas the effect is nearly neural for high levels of disturbance.

While participation in the omnivory motif is positively correlated with in-degree, a high proportion of the omnivory motif is always associated with increased persistence, while the effect of in-degree varies with disturbance. 
Thus, the effect of a high proportion of the three-species chain motif and the omnivory motif does not follow the corresponding effect of in-degree and trophic level.

@@

% The divergent effect on persistence of either having a high proportion of the three-species chain or omnivory motif can, within our Bayesian Network framework, be explained by the structure of the food web.
% When participating in a food chain motif---three-species chain or omnivory---the top species of this structure will have a higher susceptibility since disturbances amplify through the network. In our results, this is in particular evident for the three-species chain motif at higher levels of disturbance, but this also applies to species with a high proportion of the omnivory motif. While the relationship between omnivory and persistence is strongly positive at low bottom-level disturbance, the trend is reduced at high levels of disturbance. 
% This is because species participating to a high extent in omnivory motifs share the sensitivity of species participating to a high extent in three-species chain motif, and are similarly negatively affected by bottom-level disturbances. However, the top species in an omnivory motif also have the benefit of an additional prey species, located further down in the food web. This prey item can assist as a buffer, decreasing the negative effect of disturbance on the consumer species. Clearly, this additional prey is of importance, since it is more beneficial with a higher proportion of the omnivory motif for an individual species than a high proportion of the three-species chain motif. 
% Thus, the observed benefit of a high proportion of the omnivory motif can be compared to omnivores in a network feeding adaptively. Studies have demonstrated that adaptive omnivory can increase the permanence of three-species chains \citep{Fagan1997, KrivanDiehl2005, AbramsFung2010}.
% Although species in our model cannot swap prey dependent on bottom-level disturbances or lower resource densities, a species with a high proportion of omnivory will in a sense spread its risk between prey items at different trophic levels.

@@

 \cite{Cirtwill2021_inprep} analysed how species time to extinction varied depending on their network profile. They found that a high proportion of the omnivory motif resulted in a shorter time to extinction, while the other three stable motifs were associated with longer times to extinction. However, the study by Cirtwill and Wootton used a different approach using a dynamic simulation model also including top-down effects.

    % % We can move this to the SI or delete entirely.
    % Although we performed analysis of both the effect of the absolute counts of motifs and the normalized motif role on persistence, we here focus on the normalized motif role.
    % This is because, the consistency of the relationships between persistence and the number of times a species appears in any of the possible motifs suggests that the observed effects of non-normalized motif roles are actually due to degree.
    % Species with higher degrees will with high probability appear in more motifs than species with low degrees.
    % It is possible for a species with low degree  to participate in many motifs if it interacts with high-degree species (thereby obtaining many indirect interaction partners).

    % When we consider the proportion of a species' role made up by each motif, the influence of degree is largely removed.
    % Given the similarly high $R^2$ values between the models using proportions of motifs and absolute counts, it does not appear that we lose much explanatory power by removing degree. 
    % Therefore, our main analysis focus on the effect of the normalized motif role on persistence.
